---
title: "Temperature - IB"
author: "Indigo Bannister"
date: "4/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(ggplot2)
library(tidyverse)
library(janitor)
library(Kendall)

```


```{r}
# Read in data
clim <- read.table("2532217.csv",fill=TRUE,sep=',',header=T) %>% 
  mutate(date = as.Date(DATE)) %>% 
  clean_names %>% 
  select(tmax:date_2, -tobs)

```

```{r}
# filter climate data and summarize by year
clim_mwy = clim %>% 
  drop_na() %>% 
  mutate( year = year(date_2)) %>% 
  filter(year != 2021) %>% 
  group_by(year) %>% 
  summarize(tmax=mean(tmax), tmin=mean(tmin))

# Plot trends with smoothed trend line
colors <- c("Minimum Temperature" = "blue", "Maximum Temperature" = "red")

ggplot(clim_mwy, aes(x = year, y = tmax)) +
  geom_point(col = "red") +
  stat_smooth(method="lm", col="red") +
  scale_y_continuous(limits=c(min(clim_mwy$tmax), max(clim_mwy$tmax)))+
  labs(y = "Average Annual Temperature (Farenheit)",
       x = "Year",
       subtitle = "Average Daily Maximum") +
  annotate(geom="text", x=2010, y=62, label="+0.07 F/year",
              color="red") +
  theme_minimal()

ggsave("temp_max.jpeg")

ggplot(clim_mwy) +
  scale_y_continuous(limits=c(min(clim_mwy$tmin), max(clim_mwy$tmin)))+
  geom_point(data=clim_mwy, aes(x = year, tmin), col = "blue") +
  stat_smooth(data=clim_mwy, aes(x = year, tmin), col="blue", method="lm") +
  labs(y = "Average Annual Temperature (Farenheit)",
       x = "Year",
       subtitle = "Average Daily Minimum") +
  annotate(geom="text", x=2010, y=32, label="+0.05 F/year",
              color="blue") +
  theme_minimal()

ggsave("temp_min.jpeg")

```

```{r}
# Calculate slope of trend lines - tmin
res_min = lm(tmin~year, data=clim_mwy)
summary(res_min)
confint(res_min,"year", level=0.95)
predict(res_min)

# Calculate slope of trend lines - tmax
res_max = lm(tmax~year, data=clim_mwy)
summary(res_max)
confint(res_max,"year", level=0.95)
predict(res_max)

```

Slope of minimum temperature is 0.05 F/year and is statistically significant, 95% confidence interval = [0.02, 0.08]
Slope of maximum temperature is 0.07 F/year and is statistically significant, 95% confidence interval = [0.04, 0.11]

```{r}
#Run some statistical tests

MannKendall(clim_mwy$tmin)
MannKendall(clim_mwy$tmax)

```
Both show statistically significant trends.


# Extremes

```{r}
# Create a df with number of freezing days
clim$freeze = ifelse(clim$tmin <= 32, 1, 0)
clim_pk <- clim %>% 
  drop_na() %>% 
  mutate(year = year(date_2)) %>% 
  group_by(year) %>% 
  summarize(ndayfr=sum(freeze)) %>% 
  filter(year != 1968,
         year != 2021)

# Plot dataframe
ggplot(clim_pk, aes(year, ndayfr))+
  geom_point(col = 'blue')+
  stat_smooth(data=clim_pk, aes(x = year, ndayfr), col="blue", method="lm") +
  labs(x = 'Year',
       y="Number of Freezing Days") + 
  theme_minimal() 

ggsave("freeze_days.jpeg")
```
```{r}
# Calculate slope of trend lines - freezing days
res_freez = lm(ndayfr~year, data=clim_pk)
summary(res_freez)
confint(res_freez,"year", level=0.95)
```
Slope of freezing days is -0.42 days/year but is not statistically significant, 95% confidence interval = [-0.93, 0.09]

```{r}
#Run some statistical tests
MannKendall(clim_pk$ndayfr)
```
Not statistically significant

```{r}
# Compare two time periods
res_early = lm(ndayfr~year, data=subset(clim_pk, clim_pk$year %in% c(1969:1990)))
summary(res_early)

res_late = lm(ndayfr~year, data=subset(clim_pk, clim_pk$year %in% c(2000:2020)))
summary(res_late)

```
Slope for both periods was approximately -2 days per year. Capturing droughts during both periods. Both are statistically significant declines

